\added{GrapeCPNet: A Self-supervised Point Cloud Completion Network for 3D Phenotyping of Grape Bunches}

### 2025-02-13 main.tex
Updates from Overleaf
title, highlights, rm refs

@@

% Title
\title{GrapeCPNet: A deep learning integration for grape completion and phenotyping in 3D}
\author{
    Wenli Zhang \textsuperscript{a,*},
    Chao Zheng \textsuperscript{a},
    Chenhuizi Wang \textsuperscript{a},
    Pieter M. Blok \textsuperscript{b}, \\
    Haozhou Wang \textsuperscript{b},
    Wei Guo \textsuperscript{b,*}
    \blfootnote{Corresponding author: zhangwenli@bjut.edu.cn; guowei@g.ecc.u-tokyo.ac.jp }
    \blfootnote{Email addresses: zhengchao97201@163.com (Chao Zheng), Wangchenhuizi@emails.bjut.edu.cn (Chenhuizi Wang), pieter.blok@fieldphenomics.com (Pieter M. Blok), haozhou-wang@g.ecc.u-tokyo.ac.jp (Haozhou Wang), guowei@g.ecc.u-tokyo.ac.jp (Wei Guo)}
}
\date{}

@@

@@

\section{Introduction}
Fresh grape is one of the most important agricultural products and economic crops \citep{alston_grapes_2019, venkitasamy_grapes_2019}. 
Its global production and consumption is increasing year by year. 
To better guide planting strategy and monitoring growing status, several geometric parameters like counts, length, and volume at bunch level and berry level are commonly used in actual applications. \citep{luo_prediction_2024,raihen_prediction_2024,kamangir_largescale_2024,barbole_comparative_2023}. 
Among them, the number and the radius of the berries are the most important phenotypic parameters \citep{sneha_acrescale_2024}. 
Traditional manual phenotyping of traits is labor-intensive and time-consuming, and often destructive. 
To reduce the cost and increase the efficiency, the non-destructive and high-throughput acquisition of phenotypic traits based on computer vision has attracted extensive interest in recent years \citep{li_review_2020}. 
Several simple geometry traits can be acquired from 2D images by using computer vision algorithms that use pattern recognition, machine learning, and deep learning \citep{zabawa_counting_2020}. 
Projecting the 2D image plane onto the 3D world inevitably results in information loss. 
The occlusion of berries often lead to incomplete representations of grape bunches. 
Additionally, clustered berries and their shadows  create challenges for obtaining reliable measurements of phenotypic parameters through 2D image analysis, even under controlled lightning conditions  \citep{fei_comprehensive_2022,xiang_multispace_2023}.

Advances in 3D scanning and reconstruction techniques have made it possible to retain detailed spatial information about objects, enabling the acquisition of complex 3D morphological traits in grape bunches \citep{jiang_3d_2019,ni_threedimensional_2021}. 
Researchers have developed 3D phenotyping processes for grapes by combining feature-engineered point cloud processing algorithms \citep{rose_automated_2016,rist_highprecision_2018}. 
These methods primarily rely on point cloud segmentation techniques, such as edge-based segmentation, region-growing segmentation, and sphere-fitting algorithms. 
Unfortunately, traditional point cloud processing algorithms heavily rely on prior knowledge of the target object and require extensive manual feature engineering and parameter tuning to achieve optimal results in a wide range of conditions, thereby limiting their generalization performance.

State-of-the-art deep learning approaches use a data-driven feature learning from the point cloud data automatically. 
It recently shows promising adaptability and accuracy in plant 3D phenotyping \citep{schneider_predicting_2020,scholer_automated_2015,rostami_survey_2019,du_pst_2023}. Focusing on point cloud segmentation applications, 
\citet{li_maspc_2022} proposed MASPC\_Transform to segment the stems and leaves of rose bush point cloud data to obtain 3D phenotypic parameters. 
\citet{wang_3d_2022} realized leaf instance segmentation by designing the PartNet for lettuce point cloud data. 
\citet{li_psegnet_2022} proposed PSegNet for segmenting stems and leaves of tomato, tobacco, and sorghum. 
The previous studies focused on canopy level or individual plant level analysis, and demonstrated that instance segmentation alone is insufficient on berry level analysis. 
The aggregated cluster growth \citep{du_instance_2023} causes the berries to squeezed against each other, leading to self occlusion and resulting in an incomplete point cloud of berry surfaces. 
Therefore, point cloud instance segmentation can only capture partial information rather than that of whole berries.

To address the issue of partial point clouds, and consequently an underestimation of the berry traits, point cloud shape completion is one of the most commonly used solutions.
Point cloud completion is the task of generating a complete 3D shape given partial point cloud input \citep{tesema_point_2024}. 
For instance, \citet{ge_symmetrybased_2020} assume that strawberries are symmetrical, suggesting that their shape can be completed by mirroring the existing shape, provided an appropriate symmetrical plane can be identified.
\citet{marangoz_fruit_2022} proposed an super-ellipsoids matching approach to map fruits on plants and estimate their shape.
However, these methods are limited to specific fruit characteristics and require manual design of algorithms according to the features of the fruit, lacking the potential for generalized application.
Utilizing deep learning methods, researchers have designed point cloud completion networks \citep{wang_unsupervised_2021}. 
\citet{park_deepsdf_2019} introduced DeepSDF to achieve the completion of object point cloud models. It uses a learned continuous signed distance function to represent a class of shapes in latent space and is tolerant to noisy 3D input data. 
\citet{magistri_contrastive_2022} utilized DeepSDF and its latent space in an encoder-decoder network that converted a single RGB-D frame to a 3D shape of a complete fruit. 
This encoder-decoder network was extended and optimized by \citet{blok_highthroughput_2025} for high-throughput 3D shape completion of potato tubers on a commercial harvester. Outside the agricultural domain, \citet{tang_lakenet_2022} proposed a point cloud completion network (Lake-net) for furniture model point cloud datasets. 
For grape berry application, the shape of the berry can be roughly approached by the deformation of a sphere or ellipsoid depending on the species. 
Existing datasets from other fruits cannot be applied directly, and manually creating paired training datasets is time-consuming and labor-intensive.
Therefore, there is a demand to develop an analysis pipeline, capable of completing the detailed surface of an incomplete berry point cloud while requiring less labor-intensive training data preparation.

In this study, we proposed a novel berry-level 3D grape phenotyping pipeline that will overcome the occlusion issues. 
The proposed pipeline includes: 
(1) a individual berry segmentation module that segment each berry from 3D point cloud of whole bunch; 
(2) a point cloud completion module that complete the missing part of segmented berry point cloud data through automatic training data generation; 
(3) a phenotyping module that extract berry phenotypic traits including radius and volume. 
The novelty of our study is the research of this novel pipeline and its evaluation on four different fresh grape species (Red, Green, Kyoho, and Shine Muscat Grape).

@@

\subsection{Berry Instance Segmentation}
\label{sec:22}

@@

\subsection{Self-Supervised Point Cloud Completion}

The self-supervised training-based berry point cloud completion was used to predict their complete 3D model. 
Self-supervised learning is a way of training a model without annotations by clustering the data in unsupervised manners\citep{fei_selfsupervised_2023}. 
The following section summarizes the self-supervised strategy for the paired training data. It is followed by a description of the grape completion network, GrapeCPNet (Figure~\ref{fig:raw3}).

@@

\subsubsection{Berry instance segmentation}
\label{sec:251}
In this paper, we used average precision AP as the evaluation metric for berry instance segmentation (similar to the original paper of \citep{vu_softgroup_2022}). The average APs with IoU (Intersection over Union) ranging from 50\% to 95\% in steps of 5\% were used and noted as AP@50:5:95.

@@

In this paper, we use the chamfer distance (CD-distance) and F-Score to measure the effect of berry completion (similar to the original paper of \citep{yu_pointr_2021}). 
For each sample, the chamfer distance $d_{CD}$ between the predicted point set $P$ and the true point set $G$ was calculated as shown in Equation~(\ref{eq:8})

@@

@@

\section*{Data availability}
The dataset and model weight will be available on request.

>>>>

% final or draft
\usepackage[draft]{changes}

@@

% Title
% \title{GrapeCPNet: A deep learning integration for grape completion and phenotyping in 3D}
\title{\added{GrapeCPNet: A Self-supervised Point Cloud Completion Network for 3D Phenotyping of Grape Bunches}}
\author{
    Wenli Zhang \textsuperscript{a,*},
    Chao Zheng \textsuperscript{a},
    Chenhuizi Wang \textsuperscript{a},
    Pieter M. Blok \textsuperscript{b}, \\
    Haozhou Wang \textsuperscript{b},
    Wei Guo \textsuperscript{b,*}
    \blfootnote{Corresponding author: zhangwenli@bjut.edu.cn; guowei@g.ecc.u-tokyo.ac.jp }
    \blfootnote{Email addresses: zhengchao97201@163.com (Chao Zheng), Wangchenhuizi@emails.bjut.edu.cn (Chenhuizi Wang), pieter.blok@fieldphenomics.com (Pieter M. Blok), haozhou-wang@g.ecc.u-tokyo.ac.jp (Haozhou Wang), guowei@g.ecc.u-tokyo.ac.jp (Wei Guo)}
}
\date{}

@@

\textbf{\added{Highlights}}

\begin{itemize}
%   \item Novel deep learning pipeline for automated 3D grape phenotyping.
%   \item GrapeCPNet: Self-supervised network for berry points cloud completion.
%   \item High accuracy in berry instance segmentation (99.5\% average precision).
%   \item Robust performance across four grape varieties ($R^2$ of 0.855 for radius, 0.969 for 
% volume).
%   \item Nondestructive and highthroughput analysis for viticulture applications.
  \item \added{GrapeCPNet completes 3D point clouds for segmented berry instances.}
  \item \added{GrapeCPNet is trained using self-supervision, free of manual annotations.}
  \item \added{Integrated into a deep learning pipeline for automated 3D grape bunch phenotyping.}
  \item \added{Achieved over 0.85 $R^2$ and 99.5\% average precision on four grape varieties.}
\end{itemize}

@@

\section{Introduction}
Fresh grape is one of the most important agricultural products and economic crops \citep{alston_grapes_2019}. 
Its global production and consumption is increasing year by year. 
To better guide planting strategy and monitoring growing status, several geometric parameters like counts, length, and volume at bunch level and berry level are commonly used in actual applications. \citep{luo_prediction_2024,raihen_prediction_2024}. 
Among them, the number and the radius of the berries are the most important phenotypic parameters \citep{sneha_acrescale_2024}. 
Traditional manual phenotyping of traits is labor-intensive and time-consuming, and often destructive. 
To reduce the cost and increase the efficiency, the non-destructive and high-throughput acquisition of phenotypic traits based on computer vision has attracted extensive interest in recent years. 
Several simple geometry traits can be acquired from 2D images by using computer vision algorithms that use pattern recognition, machine learning, and deep learning \citep{zabawa_counting_2020}. 
Projecting the 2D image plane onto the 3D world inevitably results in information loss. 
The occlusion of berries often lead to incomplete representations of grape bunches. 
Additionally, clustered berries and their shadows  create challenges for obtaining reliable measurements of phenotypic parameters through 2D image analysis, even under controlled lightning conditions  \citep{fei_comprehensive_2022,xiang_multispace_2023}.

Advances in 3D scanning and reconstruction techniques have made it possible to retain detailed spatial information about objects, enabling the acquisition of complex 3D morphological traits in grape bunches \citep{jiang_3d_2019,ni_threedimensional_2021}. 
Researchers have developed 3D phenotyping processes for grapes by combining feature-engineered point cloud processing algorithms \citep{rose_automated_2016,rist_highprecision_2018}. 
These methods primarily rely on point cloud segmentation techniques, such as edge-based segmentation, region-growing segmentation, and sphere-fitting algorithms. 
Unfortunately, traditional point cloud processing algorithms heavily rely on prior knowledge of the target object and require extensive manual feature engineering and parameter tuning to achieve optimal results in a wide range of conditions, thereby limiting their generalization performance.
\added{Implementing} state-of-the-art deep learning approaches\replaced{, which automatically learn features}{use a data-driven feature learning} from the point cloud data \replaced{, can  reduce the workload of designing robust algorithms for complex and diverse conditions}{ automatically}.
\deleted{It recently shows promising adaptability and accuracy in plant 3D phenotyping (schneider\_predicting\_2020, scholer\_automated\_2015). }
\deleted{Focusing on point cloud segmentation applications, }
\deleted{li\_maspc\_2022 proposed MASPC\_Transform to segment the stems and leaves of rose bush point cloud data to obtain 3D phenotypic parameters. }
\deleted{wang\_3d\_2022 realized leaf instance segmentation by designing the PartNet for lettuce point cloud data. }
\deleted{li\_psegnet\_2022 proposed PSegNet for segmenting stems and leaves of tomato, tobacco, and sorghum. }
\replaced{However, few studies have reported deep learning networks perform well in}{The previous studies focused on canopy level or individual plant level analysis, and demonstrated that instance segmentation alone is insufficient on} berry level analysis.

\added{Another challenge after the success of instance berry segmentation is the incomplete representation of berry surfaces.} 
The aggregated cluster growth causes the berries to squeezed against each other, leading to self occlusion and resulting in an incomplete point cloud of berry surfaces \citep{du_instance_2023}. 
Therefore, point cloud instance segmentation can only capture partial information rather than that of whole berries.
To address \replaced{this issue}{the issue of partial point clouds,} and consequently an underestimation of the berry traits, point cloud shape completion is one of the most commonly used solutions.
Point cloud completion is the task of generating a complete 3D shape given partial point cloud input \citep{tesema_point_2024}. 
For instance, \citet{ge_symmetrybased_2020} assume that strawberries are symmetrical, suggesting that their shape can be completed by mirroring the existing shape, provided an appropriate symmetrical plane can be identified.
\citet{marangoz_fruit_2022} proposed an super-ellipsoids matching approach to map fruits on plants and estimate their shape.
However, these methods are limited to specific fruit characteristics and require manual design of algorithms according to the features of the fruit, lacking the potential for generalized application.
Utilizing deep learning methods, researchers have designed point cloud completion networks \citep{wang_unsupervised_2021}. 
\citet{park_deepsdf_2019} introduced DeepSDF to achieve the completion of object point cloud models. It uses a learned continuous signed distance function to represent a class of shapes in latent space and is tolerant to noisy 3D input data. 
\citet{magistri_contrastive_2022} utilized DeepSDF and its latent space in an encoder-decoder network that converted a single RGB-D frame to a 3D shape of a complete fruit. 
This encoder-decoder network was extended and optimized by \citet{blok_highthroughput_2025} for high-throughput 3D shape completion of potato tubers on a commercial harvester. Outside the agricultural domain, \citet{tang_lakenet_2022} proposed a point cloud completion network (Lake-net) for furniture model point cloud datasets. 
For grape berry application, the shape of the berry can be roughly approached by the deformation of a sphere or ellipsoid depending on the species. 
Existing datasets from other fruits cannot be applied directly, and manually creating paired training datasets is time-consuming and labor-intensive.
Therefore, there is a demand to develop an analysis pipeline, capable of completing the detailed surface of an incomplete berry point cloud while requiring less labor-intensive training data preparation.

In this study, we proposed a novel berry-level 3D grape phenotyping pipeline that will overcome the occlusion issues. 
The proposed pipeline includes: 
(1) a individual berry segmentation module that segment each berry from 3D point cloud of whole bunch; 
(2) a point cloud completion module that complete the missing part of segmented berry point cloud data through automatic training data generation; 
(3) a phenotyping module that extract berry phenotypic traits including radius and volume. 
\replaced{The proposed pipeline was evaluated}{The novelty of our study is the research of this novel pipeline and its evaluation} on four different fresh grape species (Red, Green, Kyoho, and Shine Muscat Grape).

@@

\subsection{Berry Instance Segmentation \added{Module}}
\label{sec:22}

@@

\subsection{Self-Supervised Point Cloud Completion \added{Module}}

The self-supervised training-based berry point cloud completion was used to predict their complete 3D model. 
Self-supervised learning is a way of training a model without annotations by clustering the data in unsupervised manners. 
The following section summarizes the self-supervised strategy for the paired training data. It is followed by a description of the grape completion network, GrapeCPNet (Figure~\ref{fig:raw3}).

@@

\subsubsection{Berry instance segmentation}
\label{sec:251}
In this paper, we used average precision AP as the evaluation metric for berry instance segmentation, similar to the original paper of \citet{vu_softgroup_2022}. The average APs with IoU (Intersection over Union) ranging from 50\% to 95\% in steps of 5\% were used and noted as AP@50:5:95.

@@

In this paper, we use the chamfer distance (CD-distance) and F-Score to measure the effect of berry completion,  similar to the original paper of \citet{yu_pointr_2021}. 
For each sample, the chamfer distance $d_{CD}$ between the predicted point set $P$ and the true point set $G$ was calculated as shown in Equation~(\ref{eq:8})

@@

\section{Conclusions}

aaa

@@

\section*{Data availability}
The dataset and model weight for the test usage is available here: https://github.com/I3-Laboratory/GrapeCPNet. The whole dataset used in this paper is available upon request.
